# Analysis: 1 numeric by 1 factor

```{r debug, echo=FALSE}
require(ggplot2, quietly=T)
opts_chunk$set(comment="") # remove the comment symbol from knitr output
opts_chunk$set(message=F, warning=F) # unless debugging
```

## Inputs

Selected variables in dataframe mydf
* Numeric: numeric1
* Factor: factor1

### Index Plot

```{r indexplot, fig.width=7, fig.height=6}
ggplot(data=mydf, aes(x=row.names(mydf), y=numeric1, color=factor1)) +
  geom_point()
```

## Summary statistics

```{r summary, comment=NA, results="asis"}
require(plyr)
kable(ddply(mydf, "factor1", summarize, numeric1.mean = mean(numeric1), numeric1.min=min(numeric1), numeric1.max=max(numeric1)), format="html", row.names=NA, table.attr='class="data table table-bordered table-condensed"')
```

### Plot of means

```{r normality, fig.width=7, fig.height=6, warning=FALSE, message=FALSE}
with(mydf, dkPlotMeans2(numeric1, factor1))
```

## Spread of values

Visually check that the distributions are normal and show homoscedasticity (similar variance).

```{r histo, fig.width=7, fig.height=6, warning=FALSE, message=FALSE}
ggplot(mydf, aes(x=numeric1, fill=factor1)) +
  geom_histogram()
```

```{r boxplot, fig.width=7, fig.height=6, warning=FALSE, message=FALSE}
ggplot(mydf, aes(x=factor1, y=numeric1)) +
  geom_boxplot()
```

### Normality tests

If either of the numeric ranges is non-normal consider using transformations or non-parametric tests (Wilcoxon if 2 sample). If the Shapiro-Wilk test p value is < 0.05 then the distribution is not normal.

```{r}
ddply(mydf, .(factor1), summarize, shapiro.p = shapiro.test(numeric1)$p.value)
```

### Variance tests

T-test, and ANOVA require homogeneity of variances (homoscedasticity). If variances are significantly different (p < 0.05) then use Kolmogorov-Smirnov test (see below).

<% if (nlevels(mydf$factor1) == 2) { %>

```{r variancetest}
  var.test(numeric1~factor1, mydf)
```

<% } else { %>

```{r variancetest}
  bartlett.test(mydf$numeric1, mydf$factor1)
```

<% } %>

<% if (nlevels(mydf$factor1) == 2) { %>

## T-Tests

```{r ttest}
t.test(numeric1 ~ factor1, mydf)
```

### Non-parametric

```{r}
with(mydf, wilcox.test(numeric1 ~ factor1))
```

<% } %>

## Anova

```{r anova}
aov1 = aov(numeric1~factor1, data=mydf)
summary(aov1)
```

<%= ifelse(summary(aov1)$"Pr(>F)" < 0.05, 
  "Anova P is < 0.05, therefore it is acceptable to continue with post-hoc comparisons",
  "Anova P is > 0.05, there is no difference across groups, post-hoc comparisons are not valid") %>

### Non-parametric

```{r}
with(mydf, kruskal.test(numeric1~factor1))
```

### Post-hoc comparisons using Tukey HSD

```{r tukey}
thsd1 = TukeyHSD(aov1)
thsd1
```

```{r tukeyplot, fig.width=7, fig.height=6}
plot(thsd1)
```

### Post-hoc comparisons using pairwise t.test with Bonferonni p adjustment

```{r pairwisettest}
with(mydf, pairwise.t.test(numeric1, factor1, p.adj="bonf"))
```

<% if (nlevels(mydf$factor1)==2) { %>

## Difference in distribution (Kolmogorov-Smirnov)

```{r komogorov}
with(mydf, ks.test(numeric1[factor1==factor1[levels()[1]]], numeric1[factor1==factor1[levels()[2]]]))
```
<% } %>