```{r setup, echo=FALSE}
require(ggplot2, quietly=T)
require(grid, quietly=T)
opts_chunk$set(comment="") # remove the comment symbol from knitr output
opts_chunk$set(message=F, warning=F) # unless debugging
```

# Analysis: 2 Numerics

## Inputs

Selected variables in dataframe mydf
* numeric1
* numeric2

## Summary statistics

```{r summary}
summary(mydf[, c("numeric1","numeric2")])
```

## Scatter Plot

```{r scatterplot, fig.width=7, fig.height=6}
ggplot(mydf, aes(x=numeric1, y=numeric2)) +
  geom_point() +
  geom_smooth()
```

## Correlation Matrix

```{r corrplot, fig.width=7, fig.height=6}
dkPairs(mydf[,c("numeric1", "numeric2")])
```

```{r corrplot2, fig.width=7, fig.height=6}
require(ez, quietly=T) 
ezCor(mydf[,c("numeric1", "numeric2")])
```

## Correlation Tests

Correlation coefficient (measure of how linear 1.0 = straight line). Values of > 0.7 are generally considered a large effect size.

```{r pearson}
with(mydf, cor(numeric1, numeric2))
```

Test of significance of correlation coefficient: Pearson

```{r cortest}
with(mydf, cor.test(numeric1, numeric2))
```

### Reporting Correlations

<% if (cor.test(mydf$numeric1, mydf$numeric2)$p.value < 0.05) { %> 
There is a significant relationship between numeric1 and numeric2, *r=`r cor(mydf$numeric1, mydf$numeric2)`, `r dkpval(cor.test(mydf$numeric1, mydf$numeric2)$p.value)`*.
<% } else { %>
There is no significant relationship between numeric1 and numeric2 (Pearson p>0.05)
<% } %>

### R^2

R^2 indicates how much of the variation in x1 is shared (and potentially explained by) x2.

```{r r2}
with(mydf, cor(numeric1, numeric2)^2)
```

### Non-parametric tests

Test of significance: Spearman (non-parametric)

```{r spearman}
with(mydf, cor.test(numeric1, numeric2, method="spearman"))
```

Test of significance: Kendall's Tau (non-parametric). Use if small dataset with many tied ranks.

```{r kendall}
table(mydf$numeric1)[table(mydf$numeric1)>1]
table(mydf$numeric2)[table(mydf$numeric2)>1]
with(mydf, cor.test(numeric1, numeric2, method="kendall"))
```

## Regression

Regression fits a straight line to the data and finds it's slope. Regression tests check if the slope is different to 0. To interpret:
- Check that the *model* p value is < 0.05
- Examine the R squared value - this describes what proportion of the variability is explained by the model
- Check the *parameter* p values - this measures if the parameter slope is significantly different from 0

```{r regression}
mylm = lm(numeric2~numeric1, data=mydf)
summary(mylm)
```

```{r regressionplot}
ggplot(mydf, aes(x=numeric1, y=numeric2)) + 
  geom_smooth(method="lm", colour="red", size=1)  
```

### Regression Diagnostics

The following graph plots the residual values against the fitted values. The dotted line is the regression line rotated to become horizontal. The y axis measures the distance of each data point from the regression line. To meet the regression assumptions the graph should show the residuals are randomly distributed about the horizontal line:
- The spread of the residuals does not change along the x axis (homoskedasticity)
- The residuals are evenly positioned above and below the regression line along the x axis (there is no pattern to location)

```{r}
autoplot(mylm, which=1:6, mfrow=c(3,2))
```

#### Check for outliers

Are there any points that dramatically alter the 

```{r}
dfbeta(mylm)
```


## Paired T-Test

```{r}
  t.test(mydf$numeric1, mydf$numeric2, paired=T)
```

